---
title: "Connected Modules (Shinylive)"
description: "An advanced Shiny application demonstrating how multiple independent modules connect, share reactive data, and communicate within a navigation bar layout."
author: "Brian S. Yandell"
toc: true
---

[㆐ Back to Demos Gallery](index.qmd)

This displays a geyser app from
[inst/connect_modules/demo](https://github.com/byandell/geyser/tree/main/inst/connect_modules/demo),
powered directly in the page via the serverless **Shinylive** (WebAssembly).

This advanced app combines a central dataset selector module with multiple plotting and layout modules, allowing them to share state reactively. Below the application, you can view the source files for each of the 6 modules plus 1 function in separate tabs.

```{shinylive-r}
#| standalone: true
#| viewerHeight: 800
#| components: [viewer]

## file: app.R
# Load required packages for WebAssembly Shinylive pre-fetching
library(shiny)
library(bslib)
library(ggplot2)
library(dplyr)
library(DT)
library(stringr)
library(tibble)

source("datanames.R")
source("datasetsApp.R")
source("gghistApp.R")
source("ggpointApp.R")
source("histApp.R")
source("rowsApp.R")
source("switchApp.R")

# ui
ui <- bslib::page_navbar(
    title = "Geyser Modules with NavBar",
    bslib::nav_panel(
        "hist",
        histInput("hist"),
        histOutput("hist"),
        histUI("hist")
    ),
    bslib::nav_panel(
        "gghist",
        gghistInput("gghist"),
        gghistOutput("gghist"),
        gghistUI("gghist")
    ),
    bslib::nav_panel(
        "ggpoint",
        ggpointInput("ggpoint"),
        ggpointOutput("ggpoint"),
        ggpointUI("ggpoint")
    ),
    bslib::nav_panel(
        "Rows",
        shiny::titlePanel("Geyser Rows Modules"),
        rowsInput("rows"),
        rowsUI("rows")
    ),
    bslib::nav_panel(
        "Switch",
        shiny::titlePanel("Geyser Switch Modules"),
        switchInput("switch"),
        switchOutput("switch"),
        switchUI("switch")
    )
)

# server
server <- function(input, output, session) {
    histServer("hist")
    gghistServer("gghist")
    ggpointServer("ggpoint")

    rowsServer("rows")

    switchServer("switch")
}
shiny::shinyApp(ui, server)

## file: datanames.R
datanames <- function() {
    # Fallback list since help("datasets") info is not fully accessible in WebAssembly R
    dataname <- unique(c("faithful", "mtcars", "cars", "iris", "quakes", "rock", "trees"))
    classes <- sapply(dataname, function(x) {
        tryCatch(class(get(x))[1],
            error = function(e) "none"
        )
    })
    tibble::as_tibble(data.frame(name = dataname, class = classes)) |>
        dplyr::filter(classes %in% c("data.frame", "matrix"))
}

## file: datasetsApp.R
datasetsApp <- function() {
    ui <- bslib::page(
        datasetsInput("datasets"),
        datasetsUI("datasets"),
        datasetsOutput("datasets")
    )
    server <- function(input, output, session) {
        datasetsServer("datasets")
    }
    shiny::shinyApp(ui, server)
}

datasetsServer <- function(id) {
    shiny::moduleServer(id, function(input, output, session) {
        ns <- session$ns

        # Select Dataset
        data <- datanames()
        output$dataset <- shiny::renderUI({
            shiny::selectInput(ns("dataset"), "Dataset:", data$name)
        })
        dataset <- shiny::reactive({
            shiny::req(input$dataset, input$columns)
            data <- get(input$dataset)
            if (!all(input$columns %in% colnames(data))) {
                return(NULL)
            }
            data[, input$columns, drop = FALSE]
        })

        # Columns
        output$columns <- shiny::renderUI({
            choices <- shiny::req(columnnames())
            if (shiny::isTruthy(input$columns)) {
                choices <- unique(c(input$columns, choices))
            }
            shiny::selectInput(ns("columns"), "Variables:",
                choices = choices,
                selected = input$columns,
                multiple = TRUE
            )
        })
        columnnames <- shiny::reactive({
            shiny::req(input$dataset)
            names(get(input$dataset) |> as.data.frame() |>
                dplyr::select(dplyr::where(is.numeric)))
        })
        shiny::observeEvent(shiny::req(input$dataset), {
            choices <- columnnames()
            shiny::updateSelectInput(session, "columns",
                choices = choices,
                selected = NULL
            )
        })

        output$table <- DT::renderDataTable(
            shiny::req(dataset())
        )

        dataset
    })
}

datasetsInput <- function(id) {
    ns <- shiny::NS(id)
    shiny::uiOutput(ns("dataset"))
}

datasetsUI <- function(id) {
    ns <- shiny::NS(id)
    shiny::uiOutput(ns("columns"))
}

datasetsOutput <- function(id) {
    ns <- shiny::NS(id)
    DT::dataTableOutput(ns("table"))
}

## file: gghistApp.R
gghistApp <- function() {
    ui <- bslib::page(
        gghistInput("gghist"),
        gghistOutput("gghist"),
        gghistUI("gghist")
    )
    server <- function(input, output, session) {
        gghistServer("gghist")
    }
    shiny::shinyApp(ui, server)
}

gghistServer <- function(id, df = shiny::reactive(faithful)) {
    shiny::moduleServer(id, function(input, output, session) {
        ns <- session$ns

        # Output Main Plot
        output$main_plot <- shiny::renderPlot({
            shiny::req(df())
            if (ncol(df()) < 1) {
                return(ggplot2::ggplot())
            }

            xvar <- colnames(df())[1]
            p <- ggplot2::ggplot(df()) +
                ggplot2::aes(.data[[xvar]]) +
                ggplot2::geom_histogram(
                    ggplot2::aes(y = ggplot2::after_stat(density)),
                    bins = as.numeric(input$n_breaks),
                    color = "black", fill = "white"
                ) +
                ggplot2::xlab(xvar) +
                ggplot2::ggtitle(stringr::str_to_title(xvar))

            if (input$individual_obs) {
                p <- p + ggplot2::geom_rug()
            }
            if (input$density) {
                shiny::req(input$bw_adjust)
                p <- p + ggplot2::stat_density(
                    adjust = input$bw_adjust,
                    color = "blue", linewidth = 1, fill = "transparent"
                )
            }
            print(p)
        })

        # Input Bandwidth Adjustment
        output$bw_adjust <- shiny::renderUI({
            if (input$density) {
                shiny::sliderInput(
                    inputId = ns("bw_adjust"),
                    label = "Bandwidth adjustment:",
                    min = 0.2, max = 2, value = 1, step = 0.2
                )
            }
        })
    })
}

gghistInput <- function(id) {
    ns <- shiny::NS(id)
    shiny::tagList(
        shiny::selectInput(
            inputId = ns("n_breaks"),
            label = "Number of bins in histogram (approximate):",
            choices = c(10, 20, 35, 50),
            selected = 20
        ),
        shiny::checkboxInput(
            inputId = ns("individual_obs"),
            label = shiny::strong("Show individual observations"),
            value = FALSE
        ),
        shiny::checkboxInput(
            inputId = ns("density"),
            label = shiny::strong("Show density estimate"),
            value = FALSE
        )
    )
}

gghistUI <- function(id) {
    ns <- shiny::NS(id)
    shiny::uiOutput(ns("bw_adjust"))
}

gghistOutput <- function(id) {
    ns <- shiny::NS(id)
    shiny::plotOutput(ns("main_plot"), height = "300px")
}

## file: ggpointApp.R
ggpointApp <- function() {
    ui <- bslib::page(
        ggpointInput("ggpoint"),
        ggpointOutput("ggpoint"),
        ggpointUI("ggpoint")
    )
    server <- function(input, output, session) {
        ggpointServer("ggpoint")
    }
    shiny::shinyApp(ui, server)
}

ggpointServer <- function(id, df = shiny::reactive(faithful)) {
    shiny::moduleServer(id, function(input, output, session) {
        ns <- session$ns

        # Output Main Plot
        output$main_plot <- shiny::renderPlot({
            shiny::req(df())
            if (ncol(df()) < 2) {
                return(ggplot2::ggplot())
            }

            xvar <- colnames(df())[1]
            yvar <- colnames(df())[2]
            p <- ggplot2::ggplot(df()) +
                ggplot2::aes(.data[[xvar]], .data[[yvar]]) +
                ggplot2::geom_point(
                    color = "black", fill = "white"
                ) +
                ggplot2::xlab(xvar) +
                ggplot2::ylab(yvar) +
                ggplot2::ggtitle(
                    stringr::str_to_title(paste(xvar, "by", yvar))
                )

            if (input$individual_obs) {
                p <- p + ggplot2::geom_rug()
            }
            if (input$smooth) {
                shiny::req(input$bw_adjust)
                p <- p +
                    ggplot2::geom_smooth(
                        formula = "y~x", se = FALSE,
                        method = "loess", span = input$bw_adjust,
                        color = "blue", linewidth = 1
                    )
            }
            print(p)
        })

        # Input Bandwidth Adjustment
        output$bw_adjust <- shiny::renderUI({
            if (input$smooth) {
                shiny::sliderInput(
                    inputId = ns("bw_adjust"),
                    label = "Span adjustment:",
                    min = 0, max = 1, value = 0.75, step = 0.05
                )
            }
        })
    })
}

ggpointInput <- function(id) {
    ns <- shiny::NS(id)
    shiny::tagList(
        shiny::checkboxInput(
            inputId = ns("individual_obs"),
            label = shiny::strong("Show individual observations"),
            value = FALSE
        ),
        shiny::checkboxInput(
            inputId = ns("smooth"),
            label = shiny::strong("Show smooth estimate"),
            value = FALSE
        )
    )
}

ggpointUI <- function(id) {
    ns <- shiny::NS(id)
    shiny::uiOutput(ns("bw_adjust"))
}

ggpointOutput <- function(id) {
    ns <- shiny::NS(id)
    shiny::plotOutput(ns("main_plot"), height = "300px")
}

## file: histApp.R
histApp <- function() {
    ui <- bslib::page(
        histInput("hist"),
        histOutput("hist"),
        histUI("hist")
    )
    server <- function(input, output, session) {
        histServer("hist")
    }
    shiny::shinyApp(ui, server)
}

histServer <- function(id, df = shiny::reactive(faithful)) {
    shiny::moduleServer(id, function(input, output, session) {
        ns <- session$ns

        # Output Main Plot
        output$main_plot <- shiny::renderPlot({
            shiny::req(df())
            if (ncol(df()) < 1) {
                return(NULL)
            }

            xvar <- colnames(df())[1]
            graphics::hist(df()[[xvar]],
                probability = TRUE,
                breaks = as.numeric(input$n_breaks),
                xlab = xvar,
                main = stringr::str_to_title(xvar)
            )

            if (input$individual_obs) {
                graphics::rug(df()[[xvar]])
            }
            if (input$density) {
                shiny::req(df())
                shiny::req(input$bw_adjust)
                dens <- stats::density(df()[[xvar]],
                    adjust = input$bw_adjust
                )
                graphics::lines(dens, col = "blue")
            }
        })

        # Input Bandwidth Adjustment
        output$bw_adjust <- shiny::renderUI({
            if (input$density) {
                shiny::sliderInput(
                    inputId = ns("bw_adjust"),
                    label = "Bandwidth adjustment:",
                    min = 0.2, max = 2, value = 1, step = 0.2
                )
            }
        })
    })
}

histInput <- function(id) {
    ns <- shiny::NS(id)
    shiny::tagList(
        shiny::selectInput(
            inputId = ns("n_breaks"),
            label = "Number of bins in histogram (approximate):",
            choices = c(10, 20, 35, 50),
            selected = 20
        ),
        shiny::checkboxInput(
            inputId = ns("individual_obs"),
            label = shiny::strong("Show individual observations"),
            value = FALSE
        ),
        shiny::checkboxInput(
            inputId = ns("density"),
            label = shiny::strong("Show density estimate"),
            value = FALSE
        )
    )
}

histUI <- function(id) {
    ns <- shiny::NS(id)
    shiny::uiOutput(ns("bw_adjust"))
}

histOutput <- function(id) {
    ns <- shiny::NS(id)
    shiny::plotOutput(ns("main_plot"), height = "300px")
}

## file: rowsApp.R
rowsApp <- function() {
    ui <- bslib::page(
        title = "Geyser Rows Modules",
        rowsInput("rows"),
        rowsUI("rows")
    )
    server <- function(input, output, session) {
        rowsServer("rows")
    }
    shiny::shinyApp(ui, server)
}

rowsServer <- function(id) {
    shiny::moduleServer(id, function(input, output, session) {
        ns <- session$ns
        dataset <- datasetsServer("datasets")
        histServer("hist", dataset)
        gghistServer("gghist", dataset)
        ggpointServer("ggpoint", dataset)
    })
}

rowsInput <- function(id) {
    ns <- shiny::NS(id)
    bslib::layout_columns(
        datasetsInput(ns("datasets")),
        datasetsUI(ns("datasets"))
    )
}

rowsUI <- function(id) {
    ns <- shiny::NS(id)
    bslib::layout_columns(
        bslib::card(
            bslib::card_header("hist"),
            histInput(ns("hist")),
            histOutput(ns("hist")),
            histUI(ns("hist"))
        ),
        bslib::card(
            bslib::card_header("gghist"),
            gghistInput(ns("gghist")),
            gghistOutput(ns("gghist")),
            gghistUI(ns("gghist"))
        ),
        bslib::card(
            bslib::card_header("ggpoint"),
            ggpointInput(ns("ggpoint")),
            ggpointOutput(ns("ggpoint")),
            ggpointUI(ns("ggpoint"))
        )
    )
}

## file: switchApp.R
switchApp <- function() {
    ui <- bslib::page(
        switchInput("switch"),
        switchOutput("switch"),
        switchUI("switch")
    )
    server <- function(input, output, session) {
        switchServer("switch")
    }
    shiny::shinyApp(ui, server)
}

switchServer <- function(id) {
    shiny::moduleServer(id, function(input, output, session) {
        ns <- session$ns

        dataset <- datasetsServer("datasets")
        histServer("hist", dataset)
        gghistServer("gghist", dataset)
        ggpointServer("ggpoint", dataset)

        output$inputSwitch <- shiny::renderUI({
            shiny::req(input$plottype)
            get(paste0(input$plottype, "Input"))(ns(input$plottype))
        })
        output$uiSwitch <- shiny::renderUI({
            shiny::req(input$plottype)
            get(paste0(input$plottype, "UI"))(ns(input$plottype))
        })
        output$outputSwitch <- shiny::renderUI({
            shiny::req(input$plottype)
            get(paste0(input$plottype, "Output"))(ns(input$plottype))
        })
    })
}

switchInput <- function(id) {
    ns <- shiny::NS(id)
    list(
        bslib::layout_columns(
            shiny::selectInput(
                ns("plottype"), "Plot Type:",
                c("hist", "gghist", "ggpoint")
            ),
            datasetsInput(ns("datasets")),
            datasetsUI(ns("datasets"))
        ),
        shiny::uiOutput(ns("inputSwitch"))
    )
}

switchUI <- function(id) {
    ns <- shiny::NS(id)
    shiny::uiOutput(ns("uiSwitch"))
}

switchOutput <- function(id) {
    ns <- shiny::NS(id)
    shiny::uiOutput(ns("outputSwitch"))
}
```

## Source Code

Below you can view the complete source files for all modules and helper scripts that compose this application.

::: {.panel-tabset}

### app.R
```{r}
#| code: !expr readLines("../../inst/connect_modules/demo/app.R")
#| eval: false
```

### datanames.R
```{r}
#| code: !expr readLines("../../inst/connect_modules/demo/datanames.R")
#| eval: false
```

### datasetsApp.R
```{r}
#| code: !expr readLines("../../inst/connect_modules/demo/datasetsApp.R")
#| eval: false
```

### gghistApp.R
```{r}
#| code: !expr readLines("../../inst/connect_modules/demo/gghistApp.R")
#| eval: false
```

### ggpointApp.R
```{r}
#| code: !expr readLines("../../inst/connect_modules/demo/ggpointApp.R")
#| eval: false
```

### histApp.R
```{r}
#| code: !expr readLines("../../inst/connect_modules/demo/histApp.R")
#| eval: false
```

### rowsApp.R
```{r}
#| code: !expr readLines("../../inst/connect_modules/demo/rowsApp.R")
#| eval: false
```

### switchApp.R
```{r}
#| code: !expr readLines("../../inst/connect_modules/demo/switchApp.R")
#| eval: false
```

:::
