---
title: "Python Module (Shinylive)"
description: "A serverless WebAssembly-powered Python Shiny app demonstrating the implementation of custom modules in Python using shiny.module."
author: "Brian S. Yandell"
toc: true
---

[← Back to Demos Gallery](index.qmd)

This section displays a Python geyser app using
[inst/build_module/5_python/app_hist.py](https://github.com/byandell/geyser/blob/main/inst/build_module/5_python/app_hist.py),
powered directly in the page via **Shinylive** (WebAssembly).

The live application is run completely serverless in your browser. Below the application, you can view the Python source files for both the main application (`app_hist.py`) and the custom module (`geyser/hist.py`) in separate tabs.

```{shinylive-python}
#| standalone: true
#| viewerHeight: 600
#| components: [viewer]
#| requirements: [pandas, numpy, matplotlib, scipy]

## file: app.py
from shiny import App, render, ui
from geyser.hist import hist_server, hist_input, hist_output, hist_ui

app_ui = ui.page_fluid(
    hist_input("hist"),
    hist_output("hist"),
    hist_ui("hist")
)

def app_server(input, output, session):
    hist_server("hist")

app = App(app_ui, app_server)

## file: geyser/__init__.py
# Package marker

## file: geyser/io.py
# Mock io module for WebAssembly environment
def app_run(app, host="127.0.0.1", port=None):
    pass

## file: geyser/hist.py
from shiny import App, module, reactive, render, ui
import numpy as np
import pandas as pd
import matplotlib.pyplot as plt
from scipy.stats import gaussian_kde

@reactive.calc
def default_dataset():
    # In browser WebAssembly, return local dataset to prevent CORS downloads
    return pd.DataFrame({
        "duration": [3.6, 1.8, 3.333, 2.283, 4.533, 2.883, 4.7, 3.6, 1.95, 4.35, 1.833, 3.917, 4.2, 1.75, 4.7, 2.167, 1.75, 4.8, 1.6, 4.25],
        "waiting": [79, 54, 74, 62, 85, 55, 88, 85, 51, 85, 54, 84, 78, 47, 83, 52, 62, 84, 52, 79]
    })

@module.server
def hist_server(input, output, session, data_set=default_dataset):
    if data_set is None:
        return None

    @reactive.calc
    def xval():
        return data_set().columns[0]
        
    @reactive.calc
    def datacol():
        return data_set()[xval()]
    
    @render.plot
    def main_plot():
        if not isinstance(data_set(), pd.DataFrame):
            return None
        if data_set().shape[1] < 1:
            return None
        fig, ax = plt.subplots()
        n_breaks = int(input.n_breaks())
        hist_data = np.histogram(datacol(), bins=n_breaks, density=True)
        ax.bar(hist_data[1][:-1], hist_data[0], width=np.diff(hist_data[1]), edgecolor='black', align='edge')

        if input.individual_obs():
            ax.plot(datacol(), np.zeros_like(datacol()), 'r|', markersize=10)

        if input.density():
            bw_adjust = input.bw_adjust()
            kde = gaussian_kde(datacol(), bw_method=bw_adjust)
            x_grid = np.linspace(min(datacol()), max(datacol()), 1000)
            ax.plot(x_grid, kde(x_grid), color='blue')

        ax.set_xlabel(xval())
        ax.set_title(xval())
        return fig
        
    @render.ui
    def output_bw_adjust():
        if input.density():
            return ui.input_slider(
                "bw_adjust",
                "Bandwidth adjustment:",
                min=0.2,
                max=2.0,
                value=1.0,
                step=0.2
            )

@module.ui
def hist_input():
    return ui.card(
        ui.input_select(
            "n_breaks",
            "Number of bins in histogram (approximate):",
            choices=[10, 20, 35, 50],
            selected=20
        ),
        ui.input_checkbox(
            "individual_obs",
            "Show individual observations",
            value=False
        ),
        ui.input_checkbox(
            "density",
            "Show density estimate",
            value=False
        )
    )

@module.ui
def hist_output():
    return ui.output_plot("main_plot")

@module.ui
def hist_ui():
    return ui.output_ui("output_bw_adjust")
```

## Source Code

::: {.panel-tabset}

### app_hist.py
```{r}
#| code: !expr readLines("../../inst/build_module/5_python/app_hist.py")
#| eval: false
```

### hist.py (module)
```{r}
#| code: !expr readLines("../../geyser/hist.py")
#| eval: false
```

:::
